---
title: "Practical_Machine_Learning"
author: "Xiaoyu Liu"
date: "September 21, 2014"
output: html_document
---
Introduction
In nowadays, large amount of data about personal activity is collected relatively inexpensively. This report predicted how well the exercise participants' performance based on the data from accelerometers placed on the belt, forearm, arm, and dumbell of six participants.

Environment Setup
##Libraries
The following libraries were used throughout the code.
```{r}
library(caret)
library(kernlab)
library(randomForest)
library(corrplot)
```
##Testing and training Data
Two data files were provided by the course.
Loading training dataset
```{r}
# read the csv file for training 
data_training <- read.csv("./data/pml-training.csv", na.strings= c("NA",""," "))
```
Loading testing dataset
```{r}
# apply the same treatment to the final testing data
data_test <- read.csv("./data/pml-testing.csv", na.strings= c("NA",""," "))

```
#Data Cleaning
NA values in the datasets would bring negative impact on the data analysis. So they were removed from the datasets.

Clean training dataset
```{r}
# clean the data by removing columns with NAs etc
data_training_NAs <- apply(data_training, 2, function(x) {sum(is.na(x))})
data_training_clean <- data_training[,which(data_training_NAs == 0)]

# remove identifier columns such as name, timestamps etc
data_training_clean <- data_training_clean[8:length(data_training_clean)]
```
Clean testing dataset
```{r}
data_test_NAs <- apply(data_test, 2, function(x) {sum(is.na(x))})
data_test_clean <- data_test[,which(data_test_NAs == 0)]
data_test_clean <- data_test_clean[8:length(data_test_clean)]
```
Developing a prediction model
The test data set was split up into training and cross validation sets in a 70:30 ratio in order to train the model.
```{r}
# split the cleaned testing data into training and cross validation
inTrain <- createDataPartition(y = data_training_clean$classe, p = 0.7, list = FALSE)
training <- data_training_clean[inTrain, ]
crossval <- data_training_clean[-inTrain, ]
```
We use the Random Forests method to predict the classification, which applies bagging to tree learners.
```{r}
#Plot
correlMatrix <- cor(training[, -length(training)])
corrplot(correlMatrix, order = "FPC", method = "circle", type = "lower", tl.cex = 0.8,  tl.col = rgb(0, 0, 0))
```
There isn't much concern for highly correlated predictors which means that all of them can be included in the model.

A regression model was built with the outcome set to the training class and all the other variables used to predict.
```{r}
model <- randomForest(classe ~ ., data = training)
```
The model produced a very small OOB error rate of .56%. This was deemed satisfactory enough to progress the testing.

#Cross-Validataion
The model was then used to classify the remaining 30% of data. The results were placed in a confusion matrix along with the actual classifications in order to determine the accuracy of the model.

```{r}

predictCrossVal <- predict(model, crossval)
confusionMatrix(crossval$classe, predictCrossVal)

```
This model yielded a 99.3% prediction accuracy. It is robust and adequete to predict new data.

Predictions
```{r}
predictTest <- predict(model, data_test_clean)
predictTest
```
#Conclusions
It can accurately predict the exercis performance of a specific person using a relatively simple model with enough data.

